---
title: "University Student Mental Health Analysis"
author: "Xinyi Lyu"
date: "2024-02-03"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(corrplot)
library(ggplot2)
library(gridExtra)
library(sandwich)
library(lmtest)
```

# Impact of Sleeping Time and Quality on Students Stress Levels

## Data

The parameters explored in this analysis are listed in detail as follows.

stress_level: sum of Pre-PSS scores ranging from 0 to 40

Hours_sleep: number of sleeping hours on an average night scaled from 1 to 9.

Rested: encoded answers to whether participants feel well-rested after sleep, where 1 represents 'yes',
2 represents 'somewhat', and 3 represents 'no'.

sleep: The Charlotte Attitudes Towards Sleep (CATS) Scale. Scores are 
averaged for a total score between 1 to 7, with higher scores indicating more favourable attitudes 
about sleep.

Extra cleaning step: Remove 467 rows containing missing values (Python)

```{r cars}
sleeping_stress <- read.csv("cleaned_stress_data.csv")
sleeping_stress <- sleeping_stress %>% select(stress_level, Hours_sleep, Rested, sleep_attitudes)
head(sleeping_stress)
```
## Initial Exploratory/Descriptive Analysis
```{r}
summary(sleeping_stress$stress_level)
sd(sleeping_stress$stress_level)
table(sleeping_stress$stress_level)
```
```{r}
summary(sleeping_stress$Hours_sleep)
sd(sleeping_stress$Hours_sleep)
table(sleeping_stress$Hours_sleep)
```
```{r}
summary(sleeping_stress$Rested)
table(sleeping_stress$Rested)
```
```{r}
summary(sleeping_stress$sleep_attitudes)
sd(sleeping_stress$sleep_attitudes)
table(sleeping_stress$sleep_attitudes)
```
```{r}
# Calculate the correlation matrix
correlation_matrix <- cor(sleeping_stress)
corrplot(correlation_matrix, method = 'color')
correlation_matrix
```

```{r}
# Convert Rested to a factor variable
sleeping_stress$Rested_factor <- factor(sleeping_stress$Rested, levels = c(1, 2, 3), labels = c("Yes", "Somewhat", "No"))
```

```{r}
hist(data$stress_level, main = "Histogram of Stress Level", xlab = "Stress Level")

# Bar plot for Rested
ggplot(sleeping_stress, aes(x = Rested_factor, fill = Rested_factor)) +
  geom_bar() +
  labs(x = "Feeling Well-Rested", y = "Frequency") +
  ggtitle("Bar plot of Feeling Well-Rested")
```
```{r}
# Histogram of sleep_attitudes
hist(sleeping_stress$sleep_attitudes, main = "Distribution of CATS Scale Scores", xlab = "CATS Scale Scores")

# Boxplot of stress levels by CATS Scale scores
boxplot(stress_level ~ sleep_attitudes, data = sleeping_stress, main = "Stress Levels by CATS Scale Scores", xlab = "CATS Scale Scores", ylab = "Stress Level")
```

```{r}
# Scatter plot of stress_level vs. Hours_sleep
ggplot(sleeping_stress, aes(x = Hours_sleep, y = stress_level)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Hours of Sleep", y = "Stress Level") +
  ggtitle("Scatter plot of Stress Level vs. Hours of Sleep")

# Scatter plot of stress_level vs. sleep_attitudes
ggplot(sleeping_stress, aes(x = sleep_attitudes, y = stress_level)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Attitudes of Sleep", y = "Stress Level") +
  ggtitle("Scatter plot of Stress Level vs. Sleep Attitudes")

# Visualize the relationship between Rested and stress level
ggplot(sleeping_stress, aes(x = Rested_factor, y = stress_level)) +
  geom_boxplot() +
  labs(x = "Rested", y = "Stress Level") +
  ggtitle("Relationship Between Rested and Stress Level")

```

## Correlation Analysis

```{r}
# correlation coefficient between stress_level and Hours_sleep
cor.test(sleeping_stress$stress_level, sleeping_stress$Hours_sleep)
cor.test(sleeping_stress$stress_level, sleeping_stress$Hours_sleep, method = "spearman")
```
```{r}
# correlation coefficient between stress_level and sleep_attitudes
cor.test(sleeping_stress$stress_level, sleeping_stress$sleep_attitudes)
cor.test(sleeping_stress$stress_level, sleeping_stress$sleep_attitudes, method = "spearman")
```
## Analysis of Variance (ANOVA)

```{r}
anova_result <- aov(stress_level ~ Rested_factor, data = sleeping_stress)
print(summary(anova_result))
```
The categorical variable "Rested" has a highly significant effect on stress levels, as indicated by the very small p-value (2.2e-08), which is much less than the conventional significance level of 0.05.

## post hoc tests

```{r}
# Tukey HSD
tukey_test <- TukeyHSD(anova_result)
print(tukey_test)

# Perform pairwise t-tests
pairwise_t_tests <- pairwise.t.test(sleeping_stress$stress_level, sleeping_stress$Rested, p.adjust.method = "bonferroni")
print(pairwise_t_tests)
```

## Linear Regression Model

```{r}
# Fit linear regression model
lm_model <- lm(stress_level ~ Hours_sleep, data = sleeping_stress)

# Summary of the regression model
summary(lm_model)

# Check assumptions
# 1. Linearity: Assess using a scatterplot of observed vs. predicted values
plot(lm_model$fitted.values, sleeping_stress$stress_level)
abline(0, 1)

# 2. Normality of residuals: Assess using a histogram or Q-Q plot of residuals
hist(residuals(lm_model))
qqnorm(residuals(lm_model))
qqline(residuals(lm_model))

# 3. Homoscedasticity: Assess using a plot of residuals vs. fitted values
plot(lm_model$fitted.values, residuals(lm_model))

# Test regression coefficients
# Test for significance of coefficients
# Hypothesis test for Hours_sleep coefficient: H0: coefficient = 0
anova(lm_model)
```

```{r}
# Fit linear regression model with robust standard errors
lm_model_robust <- coeftest(lm_model, vcov = vcovHC(lm_model))

# Summary of the regression model with robust standard errors
lm_model_robust
```

```{r}
library(lmtest)

# Fit polynomial regression model
poly_model <- lm(stress_level ~ Hours_sleep + I(Hours_sleep^2), data = sleeping_stress)

# Summary of the polynomial regression model
summary(poly_model)

# Test for significance of coefficients
summary(poly_model)$coefficients

# Wald test for significance of coefficients
anova(lm_model, poly_model)
```
The significance of the F-statistic suggests that including the quadratic term (Hours_sleep^2) in Model 2 significantly improves the model fit compared to Model 1, which only includes the linear term. Therefore, you can conclude that the polynomial regression model with the quadratic term provides a better fit to the data than the simple linear regression model.

```{r}
# Fit cubic regression model
cubic_model <- lm(stress_level ~ Hours_sleep + I(Hours_sleep^2) + I(Hours_sleep^3), data = sleeping_stress)

# Summary of the cubic regression model
summary(cubic_model)

# Perform ANOVA to compare the cubic model with the quadratic model
anova_result <- anova(poly_model, cubic_model)

# Print ANOVA table
print(anova_result)
```
Based on the non-significant p-value, we fail to reject the null hypothesis, suggesting that adding the cubic term (Hours_sleep^3) to the model does not significantly improve the fit compared to the quadratic model. Therefore, in this case, the cubic model does not provide a significantly better fit to the data compared to the quadratic model.

### Conclusion

the polynomial regression model with the quadratic term provides a best fit to the data than the simple linear regression model.

```{r}
# Generate predicted values from the model
predicted_values <- predict(poly_model, newdata = data.frame(Hours_sleep = seq(min(sleeping_stress$Hours_sleep), max(sleeping_stress$Hours_sleep), length.out = 100)))

# Scatterplot of stress level vs. hours of sleep
ggplot(sleeping_stress, aes(x = Hours_sleep, y = stress_level)) +
  geom_point() +
  geom_line(data = data.frame(Hours_sleep = seq(min(sleeping_stress$Hours_sleep), max(sleeping_stress$Hours_sleep), length.out = 100), stress_level = predicted_values), aes(x = Hours_sleep, y = stress_level), color = "red") +
  labs(x = "Hours of Sleep", y = "Stress Level") +
  ggtitle("Scatterplot of Stress Level vs. Hours of Sleep with Polynomial Regression Model")
```

## MLR (include all four variables)

```{r}
# Fit polynomial regression model with additional variables
poly_model <- lm(stress_level ~ Hours_sleep + I(Hours_sleep^2), data = data)
multy_model <- lm(stress_level ~ Hours_sleep + I(Hours_sleep^2) + sleep_attitudes, data = data)
summary(multy_model)
anova_result <- anova(poly_model, multy_model)
print(anova_result)
```

```{r}
# Fit polynomial regression model with additional variables
multy_model_3 <- lm(stress_level ~ Hours_sleep + I(Hours_sleep^2) + sleep_attitudes + Rested, data = data)
summary(multy_model_3)
anova_result <- anova(poly_model, multy_model_3)
print(anova_result)
```

```{r}
interaction_model <- lm(stress_level ~ Hours_sleep * sleep_attitudes * Rested, data = sleeping_stress)
summary(interaction_model)
anova_result <- anova(multy_model_3, interaction_model)
print(anova_result)
```
